Water discharge quantity has a variable impact on sediment transport capacity in subglacial channels

### 2023-09-29 Delaneyetal2023b.tex
Working on the algebra

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage{placeins}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

@@

\appendix
\section{Table of variables, parameters and constants}
\begin{table}[h]
  \centering
  \caption{Variables, parameters, and constants used in this work. 
    Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier}
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

\newpage
\section{Mauro's bit about shear stress}
\label{sect:scaling}
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendices go here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}
\end{document}

>>>>

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage{placeins}
\usepackage{amsmath}

\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

@@

\appendix
\section{Table of variables, parameters and constants}
\begin{table}[h]
  \centering
  \caption{Variables, parameters, and constants used in this work.
    Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier}
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-2}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

\newpage
\section{Mauro's bit about shear stress}
\label{sect:scaling}

Here we look at three possibilities, namely, rivers, steady-state
R-channels, and pipe-flow (i.e. R-channels which do not have time to
adjust their size to discharge conditions).

Basics of hydraulics
\begin{equation}
  \Psi \propto f\frac{v^2}{D}
\end{equation}
or
\begin{equation}
  \Psi \propto f\frac{Q^2}{D^5}.
\end{equation}
Then
\begin{equation}
  \tau \propto v^2 = (Q/S)^2.
\end{equation}

For all flow types we assume that the discharge $Q$ is given.
A flow in a wide river has $\Psi$ constant (given by topography), $w \gg h$, and we assume the width can be approximated by a relation $w \propto Q^\alpha$ with $\alpha\in [0,1]$ and typically $\alpha \approx 1/2$.  A R-channel flow has $\Psi$ constant (approximated by the Shreve potential) and $S$ adjusts such that it is in steady state.  Pipe flow has $S$ fixed and $\Psi$ adjusts.

Shear stresses are given by
\begin{align}
  \text{wide river} \quad\quad &  \tau \propto \left(\frac{Q}{w}\right)^{2/3} \Psi^{2/3}\\
  \text{R-channel} \quad\quad & \tau =  \\
  \text{pipe} \quad\quad &  \tau \propto v^2 \propto\\
\end{align}

Stream power are

Basics on sediment transport
\begin{align}
  \text{MPM} \quad\quad & Q_s \propto w \tau^{3/2}\\
  \text{EH}  \quad\quad& Q_s \propto w \tau^{5/2}\\
  \text{Bagnold}  \quad\quad & Q_s \propto w (\Psi Q)^{3/2}
\end{align}
for conditions well above transport onset threshold.

% \begin{tabular}{lll}
%  & Shear stress & Stream power\\[0pt]
%  & \(\tau \propto\) & \(\Omega \propto\)\\[0pt]
% \hline
% River & \(\big(\tfrac{Q}{w}\big)^{2/3} \Psi^{2/3}\) & \(Q \Psi\)\\[3pt]
% R-channel & \(Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\
% Pipe & \(Q^2\) & \(Q^3\)\\
% \end{tabular}

\noindent
\begin{tabular}{llllll}
Kind & Fixed & Determined via D-W & Additional relations & Shear stress & Stream p.\\[0pt]
 &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\[0pt]
\hline
River & \(\Psi\) & \(H \propto (f \Psi^{-1} Q^{2-2\alpha})^{1/3}\) & \(w\propto Q^\alpha\), \(S=WH\), \(D_h=H\) & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\[0pt]
R-ch. & \(\Psi\) & \(S \propto (f \Psi^{-1} Q^2)^{2/5}\) & \(D_h\propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\[0pt]
Pipe & \(S\) & \(\Psi \propto f Q^2 S^{-5/2}\) & \(D_h\propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\[0pt]
\end{tabular}

\vspace{10mm}

\noindent
\begin{tabular}{llll}
 & MPM & EH & Bagnold\\[0pt]
\(Q_s \propto\) & \$W$\backslash$, \(\tau\)\textsuperscript{3/2} & \(W\, \tau^{5/2}\) & \(W\, \Omega^{3/2} H^{-2/3}\)\\[0pt]
\hline
River & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{13\alpha/9 - 19/18} \, \Psi^{31/18}\)\\[0pt]
R-ch. & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{1/15}\, Q^{49/30}\, \Psi^{43/30}\)\\[0pt]
Pipe & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-43/12}\)\\[0pt]
\end{tabular}\\
\mauro{That negative exponent of $f$ for Bagnold-River is pretty crazy! Not sure what that is about.}

\end{document}
